Speculation in foreign exchange: noise or information?

### 2017-11-27 Knitr/FinancialEconomics.tex
This is ready to go to Pascal I think.  Look at conclusion.,

\documentclass[preprint,12pt,authoryear]{elsarticle}
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
% \documentclass[final,1p,times,authoryear]{elsarticle}
%\documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%\documentclass[final,3p,times,authoryear]{elsarticle}
%\documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%\documentclass[final,5p,times,authoryear]{elsarticle}
%\documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}
\usepackage{rotating}
\usepackage{amsmath}
%\usepackage{setspace}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
\begin{document}

\section*{Introduction}
Does speculation facilitate price discovery or instability?  If it is orice discovery, it is beneficial and should be encouraged; if it is instability, welfare is enhanced by its reduction. This paper seeks to distinguish between these two characteristics by analysing those times when specualtion in the foreign exchange market is most extreme.  A series of event studies are conducted on the extremes of speculative activity and speculative sentiment. The finding that spculative extremes do not provide information about subsequent returns implies that specualtion is part of the process of price discovery and that efforts to reduce it would leave to a reduction in information efficiency.

If markets are effective, speculation is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. i\citet{KeynesHedge} and \citet{HicksHedge} emphasise the role that speculation plays in providing liquidity and say that there is a return to this activity. Speculators provide a service for a fee that is determined by the difference between the price that they buy and the price at which they expect to sell.  Informed speculators will tend to drive out the uninformed noise-traders.  \citet{FriedmanPositive} asserted that successful speculators would be those buying below intrinsic value and selling above that rate. Therefore, Friedman argued, the speculative process would reward those able to access or utilise information relative to the uninformed and, through a Darwinian process, informed speculation would dominate.  This view regards speculation as an key lubricant of financial markets as it provides liquidity and facilitates the process of price discovery whereby information is assimilated into price.

However, the later Keynes regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'\citet[p. 101]{Keynes1936}. Speculation was a problem that would cause booms, reversals and would discourage fundamental, long-term investment. Speculation would emerge from the evolution of collective sentiment and could create trends in financial asset prices.

This paper distinguishes between the two concepts of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to the movement of prices in the foreign exchange market\footnote{\label{FX}The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.  The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).}.  If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed their activity drives exchange rates towards equlibrium and prices should follow a random walk after these extremes.

The price action after periods of extreme speculative sentiment and extreme speculative positions are comparded to the results of a noise-trader model.  Noise-trader reversals in price are not evident.  However, there is evidence that speculation in the foreign exchange market is informed. The paper contributes to the understanding of speculation and uses a unique set of option prices to identify the intensity of speculative sentiment

@@

\citet[p.529]{BlackNoise} suggested that the interaction of informed and uninformed activity was essential to the smooth functioning of financial markets. Noise, he argued, was the  ``the arbitrary element in expectations''. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.

Uninformed activity provides a rationale and the opportunity for trading. Unless there is a difference of opinion about value, there is no reason to trade. Uninformed speculation can create space between price and fundamental value that can be exploited by informed speculators.  This space can encourage informed traders to pay the cost of acquiring information.  \citet{Grossman1980Impossibility} argue that there is equilibrium between the cost of obtaining information and the benefits that are expected to come from the use of the information.

\citet{EasleyPIN} use information as an explanation for the effect of large orders on stock prices.  The \emph{order effect} may be a consequence of the need to adjust inventories when a market-maker receives a large order; that effect may also be a function of the market-maker knowledge that large orders are more likely come from to informed traders seeking to realise their informational advantage.  \citet{EasleyPIN2} develop a technique to estimate the probability that trading is informed.  They find that the high liquidity stocks have the highest level of information and the most frequent arrival rate of informed traders but that price changes are limited by the volume of uniformed traders. Low liquidity stocks by contrast, have fewer information events but informed traders have a much larger effect on prices.

\citet{EasleyPIN2} propose to measure informed market activity with the bid-ask spread.  Large orders are assumed to be informed as participants seek to explit their informational advantage.\footnote{Market-makers use Bayes' Rule to update their fundamental valuation by using order flow to estimate the joint probability that there has been an information event and that it has been good news or bad news, the authors seek to operationalise the following model:
\begin{subequations}
\begin{equation}
b(t) = E[V_i|t] - \frac{\mu P_b(t)}{\varepsilon + \mu P_b(t)}(E[V-t|-\underline{V}])
\end{equation}
\begin{equation}
a(t) = E[V_i|t] + \frac{\mu P_g(t)}{\varepsilon + \mu P_g(t)}(\overline{V}E+ [V-t|])
\end{equation}
\end{subequations}
  where $b(t)$ is the bid price at time $t$; $V$ is the value of the security; $\mu$ is the share of informed traders in the market; $P_g(t)$ is the probability that there has been some good news and $P_b(t)$ is the probability that there has been some bad news; $\varepsilon$ is the share of uninformed traders in the market; $\underline{V}$ is the value if there is bad news and $\overline{V}$ is the value if there is good news.}

However, identifying large orders and making an assessment of the level of informed trading is much harder in an over-the-counter market like foreign exchange. While electronic order books have become more widely used, much of the market-making still takes place through bilateral agreements and information about the nature of transactions including size and origin of orders is not widely available.\footnote{\citet{Lyons1995Microstructure} and \citet{Evans2002Order} have gained access to proprietary data and have used the initiation of orders as an explanatory variable in an equation to explain the evolution of major exchange rates.  They find that $\$$ 1 billion of net US dollar purchase moves the Deutsche mark price of the US unit by 0.5 percent. They argue that the orders crystallise a range of information that is used to value currencies.}  \citet{LeeandReady} seek to use price action to infer the nature of the order that has been placed into the market.  The \emph{tick-test} has four classifications: the up-tick, the down-tick, the zero-up-tick and the zero-down-tick. A trade is an up-tick if the price is higher than the previous trade.  If the price is the same as the previous trade, if the previous trade was classified as an up-tick, this trade is a zero-up-tick. Under this system, the up-ticks are assumed to be buy orders and the down-ticks are sell orders. These estimates of the market activity can be used to estimate the level of informed trading.

Informed speculation is also central to the mechanism proposed by \citet{muth1961Rational} to facilitate \emph{rational expectations}. Those economic agents able to determine fundamental value will be able to speculate and drive prices towards equilibrium.    While this construction may be justified for the analysis of microstructure and market efficiency, it is an over-simplification that is inappropriate when seeking more understanding of speculation itself.  Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speculative activity.

Noise-trading caused by psychological biases and restricted information is associated with behavioural issues, over-confidence and the use of irrational or purely fashionable investment tools. See \citet{DeBondtOver} for an analysis of over-reaction and the \emph{representativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.  Traders without information may not make decisions independently of others. Information may be inferred (perhaps incorrectly) from the buying and selling conducted by other traders, causing positive feedback and trends.\footnote{This is an example of the beauty contest that has been presented by \citet[p.101]{Keynes1936}. The aim is to determine not the underlying value but what others perceive to be the underlying value.}

Liquidity trading suggests that institutional constraints may ensure that economic agents just execute customer orders at the current price or that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy. The institutional explanations are also associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.  See \citet{BrunnermeierLiquidity} and \citet{Gorton} for some discussion of the way that margin-calls and collateral constraints can force investment decisions.

\citet{FriedmanPositive} said that the informed market participants would drive the uninformed from the market.  However, a large proportion of uninformed traders in a market can undermine this process.  \citet{Delong1990noise} and \citet{shleifer1997limits} have presented models where the risk that noise-traders will move prices even further away from equilibrium before convergence can limit the activity of informed speculators when they are risk averse. This can be called \emph{noise-trader risk}.

\section{Measuring speculation}
The measurement of speculative intensity must identify how far speculators believe that the market will move in a particular direction. In the De-Long model, this would mean that for Equation \ref{eq:NP2}, it is the absolute value of $\rho - 1$.  In the absence of a daily survey of speculators, this study uses a unique series of option\emph{risk-reversal skew} as an objective measure of speculative sentiment.\footnote{It appears that this is the only period for which the data currently exist.  The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.  The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.  Risk-reversal skew is the relative price of currency put and call options for the same strike price.} The risk-reversal skew is the relationship between a call option and the equivalent put. The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not. While the Black-Scholes model is the basis for valuing options where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.   The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.   The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity.   This model allows the market practice of pricing options from the volatility implied by the model.  In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific foreign exchange option price given the strike and other parameters.

@@

The effect of extreme speculation on future returns is analysed with an event study.  See \citet{Dolly1933} for the original version of the event study method and \citet{FamaFisherJensenRoll} for an overview of the method used in this study. However, expected returns are identified as being zero rather than established through a \emph{market-model}.  The R package \emph{eventstudies} created by \citet{eventstudies} is used to manage the data.  Extreme measures of speculative sentiment and activity are the event and the cumulative abnormal returns for the whole of the event window (CARW) and the cumulative abnormal returns for after the event (CARA) are the exchange rates to be assessed.

@@

  Table \ref{tabref:RR1} records the result of the event studies that were carried out on extreme risk reversal.   For each exchange rate the second column distinguishes between extreme high and extreme low, the third  column identifies the number of extreme events that were recorded and the following two are the cumulative average returns for the whole event window (CARW) and the cumulative average returns for the event and the period after (CARA).  The windows were chosen to correspond to one day and one week of trading.  

@@

As speculators become more extreme in their opinion, exchange rates are driven in the direction of sentiment. Looking down the seventh column (the 4-day CARW) reveals the average cumulative return for the period from four days before the extreme speculative sentiment to the four days after the event.   The extreme highs are those points when the risk-reversal skew is at or above the $95^{th}$ percentile.  For the whole event window (CARW), extreme highs (labelled "Hi") are associated with positive returns and the extreme lows (labelled "Lo") are associated with negative returns. In most instances these are significantly different from zero. Significance is identified as those cases where more than 95\% of 1000 means calculated from random bootstrap samples taken from the events are above zero.   Returns that are statistically different from zero are identified by an asterisk.

>>>>

\documentclass[preprint,12pt,authoryear]{elsarticle}
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
% \documentclass[final,1p,times,authoryear]{elsarticle}
%\documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%\documentclass[final,3p,times,authoryear]{elsarticle}
%\documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%\documentclass[final,5p,times,authoryear]{elsarticle}
%\documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}
\usepackage{rotating}
\usepackage{amsmath}
%\usepackage{setspace}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
\begin{document}
\maketitle
\begin{abstract}
Does speculation facilitate price discovery or instability?  If it is price discovery, it is beneficial and should be encouraged; if it is instability, welfare is enhanced by its reduction. This paper seeks to distinguish between these two characteristics by analysing those times when speculation in the foreign exchange market is most extreme.  A series of event studies are conducted on the extremes of speculative activity and speculative sentiment. If speculation is noise,  extreme sentiment and extreme positions should lead to overshooting and increase risk of subsequent reversals. The finding that speculative extremes do not provide information about subsequent returns implies that speculation is part of the process of price discovery and that efforts to reduce it would leave to a reduction in information efficiency.

\end{abstract}

\section{Introduction}
If speculation is based on information it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. \citet{KeynesHedge} and \citet{HicksHedge} emphasised the role that speculation played in providing liquidity. Speculators provide a service for a fee that is determined by the return on their activity.  Informed speculators will tend to drive out the uninformed noise-traders \citet{FriedmanPositive}.  Successful speculators would be those buying below intrinsic value and selling above that rate.  However, the later Keynes regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'\citet[p. 101]{Keynes1936}.

This paper evaluates the relative importance of these two concepts of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to subsequent movement of foreign exchange prices. \footnote{\label{FX}The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.  The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).}.  If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes.

@@

\citet[p.529]{BlackNoise} argued that the interaction of informed and uninformed activity was essential to the smooth functioning of financial markets. Noise, he argued, was the  ``the arbitrary element in expectations''. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.

Uninformed activity provides a rationale and the opportunity for trading. Unless there is a difference of opinion about value, there is no reason to exchange. Uninformed speculation can create space between price and fundamental value that can be exploited by informed speculators.  This space can encourage informed traders to pay the cost of acquiring information.  \citet{Grossman1980Impossibility} argue that there is equilibrium between the cost of obtaining information and the benefits that are expected to come from the use of the information.

\section{Measuring speculation}
This study uses a unique series of option\emph{risk-reversal skew} as an objective measure of speculative sentiment.\footnote{It appears that this is the only period for which the data currently exist.  The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.  The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.  Risk-reversal skew is the relative price of currency put and call options for the same strike price.} The risk-reversal skew is the relationship between a call option and the equivalent put.

@@

The effect of extreme speculation on future returns is analysed with an event study.  See \citet{Dolly1933} for the original version of the event study method and \citet{FamaFisherJensenRoll} for an overview of the method used in this study.

@@

  Table \ref{tabref:RR1} records the result of the event studies that were carried out on extreme risk reversal.   For each exchange rate the second column distinguishes between extreme high and extreme low risk-reversals, the third  column identifies the number of extreme events that were recorded while the following two columns show the cumulative average returns for the whole event window (CARW) and the cumulative average returns for the event and the period after (CARA).  The windows were chosen to correspond to one day and one week of trading.  

@@

As speculators become more extreme in their opinion, exchange rates are driven in the direction of sentiment. Looking down the seventh column (the 4-day CARW) reveals the average cumulative return for the period from four days before the extreme speculative sentiment to the four days after the event.   The extreme highs are those points when the risk-reversal skew is at or above the $95^{th}$ percentile.  For the whole event window (CARW), extreme highs (labelled "Hi") are associated with positive returns and the extreme lows (labelled "Lo") are associated with negative returns. In most instances these are significantly different from zero.
